A Short Guide to Anyons and Modular Functors

### 2016-10-17 guide.tex
final tweaks

In three spatial dimensions, the process of winding one
particle around another, a \emph{monodromy}, is topologically trivial.
This is because the path can be deformed back to the
identity; there is no obstruction:
\begin{center}
\includegraphics[]{pic-monodromy3d.pdf}
\end{center}
The square root of this operation is a swap:
\begin{center}
\includegraphics[]{pic-swap.pdf}
\end{center}
For identical particles this is a symmetry of the system. 
%%%%%% \cggb{What does symmetry mean here? Since a swap of non-bosons will pick up a phase.}
Continuing with this line of thought 
leads to consideration of the \emph{symmetric group}
on $n$ letters, $S_n$. 
This group is generated by the $n-1$ swap operations
$s_1,...,s_{n-1}$ that obey the relations
\begin{align*}
    s_i^2 &= 1, \\
    s_i s_j &= s_j s_i \ \ \ \mbox{for}\ \ |i-j|>1,\\
    s_i s_{i+1} s_i &= s_{i+1} s_i s_{i+1} \ \ \ \mbox{for}\ \ 1\le i \le n-2.
\end{align*}
%This group acts on the statespace of $n$ particles $\F_n$
%via unitaries $U(\F_n)$:
%$$
%    \Phi : S_n \to U(\F_n).
%$$
Writing the Hilbert space of the system as $V$ 
we would then expect $S_n$ to act on this space via unitary transformations $U(V):$
$$
    \F : S_n \to U(V).
$$
% For bosons this is ... for fermions it is ...
This is the first example of the kind of functor we will be talking about. 
In this case it is a group representation; $\F$ is a homomorphism between
two groups.

@@

%By an \emph{isotopy} of $f:M\to N$ we mean a 
%map $g:M\to N$
%that is a continuos deformation of $f.$
Two maps of surfaces $f:M\to N$ and $f':M\to N$
will be called \emph{isotopic}
when one is a continuous deformation of the other.
In detail,
we have a continuously
parametrized family of maps $f_t:M\to N$ for $t\in [0, 1]$
such that $f_0=f,\ f_1=f'$
and the restriction of $f_t$ to the 
set of marked points $X\subset\partial M$ is constant:
$f_t|_{X} = f|_{X}$ for $t\in [0, 1].$
%and the restriction of $f_t$ to the boundary is constant:
%$f_t|_{\partial M} = f|_{\partial M}$ for $t\in [0, 1].$
Such a family $\{f_t\}_{t\in [0,1]}$ is called an \emph{isotopy}
of $f.$
This is an equivalence relation on maps $M\to N$, and the equivalence
class of $f$ under isotopy is called the \emph{isotopy class} of $f.$
%\cggb{I didn't quite see why this more formal definition is
%the same as equivalent up to continuous deformation. Why can
%I restrict to maps where the base points are fixed along the whole path?}
The weaker notion of \emph{homotopy} of maps will not be used
here, but for the maps we use
it turns out that homotopy is equivalent to isotopy.
Furthermore, we can weaken the requirement that maps be differentiable,
because every continuous map $f:M\to N$ is (continuously)
isotopic to a differentiable map~\cite{Farb2011}.

>>>>

In three spatial dimensions, the process of winding one
particle around another, a \emph{monodromy}, is topologically trivial.
This is because the path can be deformed back to the
identity; there is no obstruction:
\begin{center}
\includegraphics[]{pic-monodromy3d.pdf}
\end{center}
The square root of this operation is a swap:
\begin{center}
\includegraphics[]{pic-swap.pdf}
\end{center}
For identical particles this is a \emph{symmetry} of the system. 
%%%%%% \cggb{What does symmetry mean here? Since a swap of non-bosons will pick up a phase.}
Continuing with this line of thought 
leads to consideration of the \emph{symmetric group}
on $n$ letters, $S_n$. 
This group is generated by the $n-1$ swap operations
$s_1,...,s_{n-1}$ that obey the relations
\begin{align*}
    s_i^2 &= 1, \\
    s_i s_j &= s_j s_i \ \ \ \mbox{for}\ \ |i-j|>1,\\
    s_i s_{i+1} s_i &= s_{i+1} s_i s_{i+1} \ \ \ \mbox{for}\ \ 1\le i \le n-2.
\end{align*}
%This group acts on the statespace of $n$ particles $\F_n$
%via unitaries $U(\F_n)$:
%$$
%    \Phi : S_n \to U(\F_n).
%$$
Writing the Hilbert space of the system as $V$ 
we would then expect $S_n$ to act on this space via unitary transformations $U(V):$
$$
    \F : S_n \to U(V).
$$
% For bosons this is ... for fermions it is ...
This is the first example of the kind of functor we will be talking about. 
In this case it is a group representation; $\F$ is a homomorphism between
two groups.

@@

%By an \emph{isotopy} of $f:M\to N$ we mean a 
%map $g:M\to N$
%that is a continuos deformation of $f.$
Two maps of surfaces $f:M\to N$ and $f':M\to N$
will be called \emph{isotopic}
when one is a continuous deformation of the other.
In detail,
we have a continuously
parametrized family of maps $f_t:M\to N$ for $t\in [0, 1]$
such that $f_0=f,\ f_1=f'$
and the restriction of $f_t$ to the 
set of marked points $X\subset\partial M$ is constant:
$f_t|_{X} = f|_{X}$ for $t\in [0, 1].$
%and the restriction of $f_t$ to the boundary is constant:
%$f_t|_{\partial M} = f|_{\partial M}$ for $t\in [0, 1].$
Such a family $\{f_t\}_{t\in [0,1]}$ is called an \emph{isotopy}
of $f.$
This is an equivalence relation on maps $M\to N$, and the equivalence
class of $f$ under isotopy is called the \emph{isotopy class} of $f.$
%\cggb{I didn't quite see why this more formal definition is
%the same as equivalent up to continuous deformation. Why can
%I restrict to maps where the base points are fixed along the whole path?}
The weaker notion of homotopy of maps will not be used
here, but for the maps we use
it turns out that homotopy is equivalent to isotopy.
Furthermore, we can weaken the requirement that maps be differentiable,
because every continuous map $f:M\to N$ is (continuously)
isotopic to a differentiable map~\cite{Farb2011}.
